XS-Face: Merge Raw Data

### 2020-04-10 writeup-paper/old-drafts/draft_1_from_CogSci.Rmd
add writeup folder for apa draft

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(egg) # plot arranging
library(lme4)
library(gridExtra)
library(feather)
library(assertthat)
```
# Introduction
From early infancy, children are deeply engaged in learning from others. Even newborns tend to prefer to look at faces with direct vs. averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010development]. Further, when infants attend to video stimuli, they tend to look mostly at faces at the expense of other visual information -- though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring].

Then, however, their view of the world radically changes [@adolph2006motor]. Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently. These motor changes have significant consequences for what children see. For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].

These motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities in various ways. For example, postural changes impact how children interact with their mothers: walking (vs. crawling) infants make different kinds of object-related bids for attention from their mothers, tend to hear more action directed statements (e.g., “open it”), and tend to explore their surroundings more [@karasik2014; @thurman2017spatial]. In an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies. Thus, children’s ability to independently locomote may change their ability to access social information (e.g., faces, gaze) and in turn accelerate their learning.

Recent technological developments allow for testing of this hypothesis by documenting the experiences of infants and children from their own perspective. By using head-mounted cameras, researchers have begun to record the visual experiences of infants and children -- which even for walking children are extremely different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008]. Children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], and computational and empirical work suggests that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2013statistical]. This perspective changes over the first two years of life, as visual input transitions from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].

Here, we directly examine whether postural and locomotor developments change the availability of social information -- the presence of faces and hands. We recorded the visual experience of a group of infants in three age ranges (8, 12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated. Recent work directly examined how postural developments change access to social information in 12 month-olds in a similar paradigm, finding support for the motoric hypothesis [@franchak2017see]. Here we replicate and extend their main finding using novel computer vision techniques to allow for video analysis at scale. In particular, we capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers. We predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape, thus providing children with greater access to the social information in their environment.

# Methods
### Participants
```{r echo=FALSE, include=FALSE}
all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
  rename(subid = subID)
inc_subs <- read.csv("../data/demographics/demographics.csv")
all_subs$inc <- all_subs$reason=="included"

@@

### Procedure
All parents signed consent documents while children were fitted with the headcam. If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam. When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session. Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys. The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.

@@

### Posture and Orientation Annotation
We created custom annotations to describe the child's physical posture (e.g. standing) and the orientation of the child relative to the caregiver (e.g. far away). The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing. The caregiver's orientation was characterized as being close, far, or behind the child (independent of distance). For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child. All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012toward]. Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.

## Face and Hand Detection
We used three face detection systems to measure infants' access to faces. The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones. We used this algorithm as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual]. We next tested the performance of two face detectors that both made use of recently developed Convolutional Neural Networks (CNNs) to extract face information. The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts. For the second algorithm [OpenPose; @cao2017realtime], we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.

The OpenPose detector also provided us with the location of an agent's wrists, which we used as a proxy for hands for two reasons. First, as we did not capture children's entire visual field, the presence of a wrist is likely often indicative of the presence of a hand within the field of view. Second, hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction. 
\begin{figure*}
\centering
\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.}
\end{figure*}

@@

### Detector evaluation
To evaluate face detector performance, we hand-labeled a "gold set" of labeled frames. To account for the relatively rare appearance of faces in the dataset, we hand-labeled two types of samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames. Each sample was comprised of an equal number of frames taken from each child's video. For wrist detections, the "gold set" was constructed in the same manner, except frames with a high density of wrists came only from detections made by OpenPose.  Faces were classified as present if at least half of the face was showing; wrists were classified as present if any part of the wrist was showing.  Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 1.

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN DETECTIONS
d <- read_csv("../data/consolidateddata/consolidated_data_4detectors_march7.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

@@

# Results
<!-- First, we report developmental shifts in infants' posture and their orientation relative to their caregiver. Then, we explore how these changes influence children's visual access to faces and wrists/hands across this developmental time range. Finally, we explore how these changes impact the accessibility of faces and wrists/hands during labeling events. -->

@@

```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
detectionsByAge <- dlong %>%
  filter(detectorType!="OpenPose-Faces") %>% # for brevity
  filter(detectorType!="ViolaJones-Faces") %>% # low performance
  filter(!is.na(age.at.test)) %>%
  group_by(subid, age.at.test, detectorType) %>%
  summarise(detections = mean(detection)) %>%
  left_join(inc_subs) %>%
  mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video

ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
  geom_jitter(width = .25) +
  geom_smooth(size=1) +
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
  labs(y = "Prop. detected", x = element_blank()) +
  theme_few() + 
  theme(legend.position="none") +
  facet_grid(~detectorType) 
```

@@

```{r echo=FALSE, include=FALSE }
source("helper.R")

aroundNamings_Posture_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceMT) %>% ## for covergence with helper.R file, which looks for this variable name
  do(summarize.naming(.)) %>%
  group_by(subid, posture) %>%
  summarize(propFacesDetected = mean(detections)) %>%
  filter(!is.na(propFacesDetected)) %>%
  filter(!is.na(posture)) %>%
  group_by(posture) %>%
  multi_boot_standard(col = "propFacesDetected")

aroundNamings_Orient_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceMT) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, orientation) %>%
  summarize(propFacesDetected = mean(detections)) %>%
  filter(!is.na(propFacesDetected)) %>%
  filter(!is.na(orientation)) %>%
  group_by(orientation) %>%
  multi_boot_standard(col = "propFacesDetected")

###
aroundNamings_Posture_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, posture) %>%
  summarize(propWristsDetected = mean(detections)) %>%
  filter(!is.na(propWristsDetected)) %>%
  filter(!is.na(posture)) %>%
  group_by(posture) %>%
  multi_boot_standard(col = "propWristsDetected")

aroundNamings_Orient_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, orientation) %>%
  summarize(propWristsDetected = mean(detections)) %>%
  filter(!is.na(propWristsDetected)) %>%
  filter(!is.na(orientation)) %>%
  group_by(orientation) %>%
  multi_boot_standard(col = "propWristsDetected") 
```

>>>>

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(egg) # plot arranging
library(lme4)
library(gridExtra)
library(feather)
library(assertthat)
source("helper.R") ## script for merging in naming annotations.
```
# Introduction
# Methods
### Participants
```{r echo=FALSE, include=FALSE}
all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
  rename(subid = subID)
inc_subs <- read.csv("../data/demographics/demographics.csv")
all_subs$inc <- all_subs$reason=="included"

@@

### Procedure
All parents signed consent documents while children were fitted with the headcam. If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam. When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session. Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." These three pairs were taken from a superset of five paired groups of novel and familiar objects (familiar objects: cat, toothbrush, car, rubber duck, toy car; all stimuli avaliable at XXX. All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys. The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.

@@

### Posture and Orientation Annotation
We created custom annotations to describe the child's physical posture (e.g. standing) and the orientation of the child relative to the caregiver (e.g. far away). The child's posture was categorized as being held/carried by the parent, prone (crawling or lying on back or front), sitting, or standing. The caregiver's orientation was characterized as being far away from the child (out of reaching distance), behind the child (and within reaching distance) or close to the child (within reaching distnace either to the front or the side of the child).  Activities were coded in 2-3 second windows, meaning that if a child briefly sat for 1 second and then went back to standing, this brief transition was ignored. All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012toward]. Times when the child was out of view of the tripod camera or when children's action were highly sporadic  were marked as uncodable and were excluded from both of these these annotations.

### Naming annotations
Naming events were coded by a trained coder who listened  to the audio from the head-mounted camera video (using the Transcriptions software for MacOS). These naming events were coded relative to word onset for each of eight possible objects (ball, cat, brush,  car,  duck, zem, gimo,  tima, manu, and gazzer) and only proper names were coded (e.g., "brush" and zem" were both coded, "the yellow one" was not). If a parent used a variant of the object name (e.g., "kitty" instead of "cat"), these naming events were recorded as the standard name for the object. Timestamps from these manual codings were automatically synced to the timestamps for each frame in each video. Full transcriptions are hosted in the publicly avaliable repository for this manuscript.

## Face and Hand Detection
We used three face detection systems to measure infants' access to faces. The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones. We used this algorithm as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual] which are likely to be common in video data from these head-mounted cameras. We next tested the performance of two face detectors that both make use of recently developed Convolutional Neural Networks (CNNs) to extract face information. The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts.

We used the outputs of the second algorithm [OpenPose; @cao2017realtime] to create proxies for both face and hand detection. As OpenPose seeks to estimate a body frame and then derive estimates of the face/hand positions from this body frame, we reasoned that it might do a better job of detection face/hand information in situations when part of the hand/face are not visible given that  the camera does not entirely capture the child's field of view. Thus, for face detection we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose. The OpenPose detector also provided us with the location of an agent's wrists, which we used as a proxy for hands for two reasons. First, as we did not capture children's entire visual field, the presence of a wrist is likely often indicative of the presence of a hand within the field of view. Second, hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.

\begin{figure*}
\centering
\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.}
\end{figure*}

@@

### Detector evaluation
To evaluate face detector performance, we hand-labeled a "gold set" of labeled frames (using RectLabel software for MacOS). To account for the relatively rare appearance of faces in the dataset, we hand-labeled two types of samples: a sample containing a high density of faces as reported by our two new detectors (half detections by MTCNN, half detections by OpenPose) and a random sample from the remaining frames. Each sample was comprised of an equal number of frames taken from each child's video, comprising roughly 1000 frames overall (14 frames for each sample type for each of 36 children). For wrist detections, the "gold set" was constructed in the same manner, except frames with a high density of wrists came only from detections made by OpenPose. For hand-annotation, faces were classified as present if at least half of the face was showing; while wrists were classified as present if any part of the wrist was showing.  Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 1.

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN DETECTIONS
d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_march7.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation)) %>%
  select(-handOP)

@@

# Results
First, we report developmental shifts in infants' posture and their orientation relative to their caregiver. Then, we explore how these changes influence children's visual access to faces and wrists/hands across this developmental time range. Finally, we explore how these changes impact the accessibility of faces and wrists/hands during labeling events.

@@

### extra analyses not included right now
```{r facesAndHands, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "."}
facesAndHands <- d %>%
  mutate(faceAndHand = faceMT & wristOP) %>%
  filter(!is.na(age.at.test)) %>%
  filter(!is.na(posture)) %>%
  group_by(subid, age.at.test, posture) %>%
  summarise(faceDetect = sum(faceMT), faceAndHand = sum(faceAndHand), propBoth =faceAndHand /  faceDetect) %>%
  left_join(inc_subs) 
  # mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video

ggplot(facesAndHands, aes(x = factor(age.grp), y = propBoth, col = posture)) + 
  geom_boxplot(outlier.size = 0.5) + 
  labs(y = "Log face size", x = element_blank()) +
  theme_few() +
  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
  scale_y_continuous(breaks=seq(0,.60,.20))

# 
# ggplot(facesAndHands, aes(x = age.at.test, y = propBoth, size=len.min)) +
#   geom_jitter(width = .25) +
#   geom_smooth(size=1) +
#   scale_x_continuous(breaks=c(8,12,16)) +
#   scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
#   labs(y = "Prop. both faces & hands", x = element_blank()) +
#   theme_few() + 
#   theme(legend.position="none") 
```

```{r faceSizeByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap ="}
faceSizeByAge <- dlong %>%
  filter(detectorType!="OpenPose-Faces") %>% # for brevity
  filter(detectorType!="ViolaJones-Faces") %>% # low performance
  filter(detection == TRUE) %>%
  filter(!is.na(age.at.test)) %>%
  mutate(logFaceSize = log(w*h)) %>%
  group_by(subid, age.at.test, detectorType) %>%
  summarise(logFaceSize = mean(logFaceSize)) %>%
  left_join(inc_subs) %>%
  mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video

ggplot(faceSizeByAge, aes(x = age.at.test, y = logFaceSize, color = detectorType, size=len.min)) +
  geom_jitter(width = .25) +
  geom_smooth(size=1) +
  scale_x_continuous(breaks=c(8,12,16)) +
  # scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
  labs(y = "Face size", x = element_blank()) +
  theme_few() + 
  theme(legend.position="none") +
  facet_grid(~detectorType) 
```

```{r faceSizeByPosture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "."}
faceSizeByPosture <- dlong %>%
  filter(detectorType=="MTCNN-Faces") %>% # has bounding box
  filter(detection == TRUE) %>%
  filter(!is.na(age.at.test)) %>%
  filter(!is.na(posture)) %>%
  mutate(logFaceSize = log(w*h)) %>%
  group_by(subid, age.at.test, posture) %>%
  summarise(logFaceSize = mean(logFaceSize), meanVert = mean(y)) %>%
  left_join(inc_subs)

ggplot(faceSizeByPosture, aes(x = factor(age.grp), y = logFaceSize, col = posture)) + 
  geom_boxplot(outlier.size = 0.5) + 
  labs(y = "Log face size", x = element_blank()) +
  theme_few() +
  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
  scale_y_continuous(breaks=seq(0,.60,.20))

```{r faceSizeByOrientation}

faceSizeByOrientation <- dlong %>%
  filter(detectorType=="MTCNN-Faces") %>% # has bounding box
  filter(detection == TRUE) %>%
  filter(!is.na(age.at.test)) %>%
  filter(!is.na(orientation)) %>%
  mutate(logFaceSize = log(w*h)) %>%
  group_by(subid, age.at.test, orientation) %>%
  summarise(logFaceSize = mean(logFaceSize), meanVert = mean(y)) %>%
  left_join(inc_subs)

ggplot(faceSizeByOrientation, aes(x = factor(age.grp), y = logFaceSize, col = orientation)) + 
  geom_boxplot(outlier.size = 0.5) + 
  labs(y = "Log face size", x = element_blank()) +
  theme_few() +
  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
  scale_y_continuous(breaks=seq(0,.60,.20))
```

@@

```{r}
#### Face detections during naming events vs. baseline by posture

aroundNamings_Posture_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceMT) %>% ## for covergence with helper.R file, which looks for this variable name
  do(summarize.naming(.)) %>%
  group_by(subid, posture) %>%
  summarize(propFacesDetected = mean(detections)) %>%
  filter(!is.na(propFacesDetected)) %>%
  filter(!is.na(posture)) %>%
  group_by(posture) %>%
  multi_boot_standard(col = "propFacesDetected")  %>%
  mutate(event_type = "naming")

Posture_Faces_Baseline <- d %>%
  group_by(subid) %>%
  mutate(detections = faceMT) %>%
  group_by(subid, posture) %>%
  summarize(propFacesDetected = mean(detections)) %>%
  filter(!is.na(propFacesDetected)) %>%
  filter(!is.na(posture)) %>%
  group_by(posture) %>%
  multi_boot_standard(col = "propFacesDetected") %>%
  mutate(event_type = "baseline")

namingVsBaseline_PostueFaces <- Posture_Faces_Baseline %>%
  full_join(aroundNamings_Posture_Faces)

ggplot(namingVsBaseline_PostueFaces, aes(x = posture, y = mean, col = event_type)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. faces") +
  xlab("") +
  theme_few() +
  theme(legend.position="left")  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))
```

```{r}
#### Face detections during naming events vs. baseline by orientation
aroundNamings_Orient_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceMT) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, orientation) %>%
  summarize(propFacesDetected = mean(detections)) %>%
  filter(!is.na(propFacesDetected)) %>%
  filter(!is.na(orientation)) %>%
  group_by(orientation) %>%
  multi_boot_standard(col = "propFacesDetected") %>%
  mutate(event_type = "naming")

Orient_Faces_Baseline <- d %>%
  group_by(subid) %>%
  mutate(detections = faceMT) %>%
  group_by(subid, orientation) %>%
  summarize(propFacesDetected = mean(detections)) %>%
  filter(!is.na(propFacesDetected)) %>%
  filter(!is.na(orientation)) %>%
  group_by(orientation) %>%
  multi_boot_standard(col = "propFacesDetected") %>%
  mutate(event_type = "baseline")

###
namingVsBaseline_Orientation_Faces <- Orient_Faces_Baseline %>%
  full_join(aroundNamings_Orient_Faces)

ggplot(namingVsBaseline_Orientation_Faces, aes(x = orientation, y = mean, col = event_type)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. faces") +
  xlab("") +
  theme_few() +
  theme(legend.position="left")  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))

```{r}
#### Wrist detections during naming events vs. baseline by posture
aroundNamings_Posture_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, posture) %>%
  summarize(propWristsDetected = mean(detections)) %>%
  filter(!is.na(propWristsDetected)) %>%
  filter(!is.na(posture)) %>%
  group_by(posture) %>%
  multi_boot_standard(col = "propWristsDetected")  %>%
  mutate(event_type = "naming")

Posture_Wrists_Baseline <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  group_by(subid, posture) %>%
  summarize(propWristsDetected = mean(detections)) %>%
  filter(!is.na(propWristsDetected)) %>%
  filter(!is.na(posture)) %>%
  group_by(posture) %>%
  multi_boot_standard(col = "propWristsDetected") %>%
  mutate(event_type = "baseline")

namingVsBaseline_Posture_Wrists <- Posture_Wrists_Baseline %>%
  full_join(aroundNamings_Posture_Wrists)

ggplot(namingVsBaseline_Posture_Wrists, aes(x = posture, y = mean, col = event_type)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. wrists") +
  xlab("") +
  theme_few() +
  # theme(legend.position="left")  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))
```

```{r namingVBasline_Orientation_Wrists}
#### Wrist detections during naming events vs. baseline by orientation
aroundNamings_Orient_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, orientation) %>%
  summarize(propWristsDetected = mean(detections)) %>%
  filter(!is.na(propWristsDetected)) %>%
  filter(!is.na(orientation)) %>%
  group_by(orientation) %>%
  multi_boot_standard(col = "propWristsDetected") %>%
  mutate(event_type = "naming")

Orient_Wrists_Baseline <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  group_by(subid, orientation) %>%
  summarize(propWristsDetected = mean(detections)) %>%
  filter(!is.na(propWristsDetected)) %>%
  filter(!is.na(orientation)) %>%
  group_by(orientation) %>%
  multi_boot_standard(col = "propWristsDetected") %>%
  mutate(event_type = "baseline")

namingVsBaseline_Orient_Wrists <- Orient_Wrists_Baseline %>%
  full_join(aroundNamings_Orient_Wrists)

p2=ggplot(namingVsBaseline_Orient_Wrists, aes(x = orientation, y = mean, col = event_type)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. wrists") +
  xlab("") +
  theme_few() +
  theme(legend.position="left")  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))

### 2020-04-10 writeup-paper/old-drafts/draft_march2020.Rmd
add writeup folder for apa draft

abstract: "The ability to process social information is a critical component of children's early language and cognitive development. However, as children reach their first birthday, they begin to locomote themselves, dramatically affecting their visual access to this information. How do these postural and locomotor changes affect children's access to the social information relevant for word-learning? Here, we explore this question by using head-mounted cameras to record 36 infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to estimate the proportion of faces and hands in infants' environments. We find that infants' posture and orientation to their caregiver modulates their access to social information, confirming previous work that suggests motoric developments play a significant role in the emergence of children's linguistic and social capacities. We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience."

@@

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(egg) # plot arranging
library(lme4)
library(gridExtra)
library(feather)
library(assertthat)
```
# Introduction
From early infancy, children are deeply engaged in learning from others. Even newborns tend to prefer to look at faces with direct vs. averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010development]. Further, when infants attend to video stimuli, they tend to look mostly at faces at the expense of other visual information -- though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring].

Then, however, their view of the world radically changes [@adolph2006motor]. Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently. These motor changes have significant consequences for what children see. For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].

@@

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN DETECTIONS
d <- read_csv("../data/consolidateddata/consolidated_data_4detectors_march7.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

>>>>

abstract: "The ability to process social information is a critical component of children's early language and cognitive development. However, as children reach their first birthday, they begin to locomote themselves, dramatically affecting their visual access to this information. How do these postural and locomotor changes affect children's access to the social information relevant for word-learning? Here, we explore this question by using head-mounted cameras to record 36 infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to estimate the proportion of faces and hands in infants' environments. We find that infants' posture and orientation to their caregiver modulates their access to social information, confirming previous work that suggests motoric developments play a significant role in the emergence of children's linguistic and social capacities. We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience as they change over development."

@@

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(egg) # plot arranging
library(lme4)
library(gridExtra)
library(feather)
library(assertthat)
```

# Introduction
Children are remarkably skilled language learners, connecting arbitrary labels (“cup”) with specific visual concepts at a rapid pace through the first two years of life. Yet children do not learn words in a vacuum but rather in a rich social environment, where social cues provided by speakers (e.g., eye-gaze) provide strong scaffolding for this learning process. Indeed, some empirical work suggests children’s ability to effectively process these social cues may be a key factor in their early language development.  In a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies [@carpenter1998]. Furthermore, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months [@brooks2005] and throughout the second year of life [@brooks2008].

Around their first birthday, however, their view of the world radically changes [@adolph2006motor] as their motor abilities improve dramatically, allowing them to locomote independently for the first time.  These motor changes have significant consequences for what children see. For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].

@@

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN DETECTIONS
d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_march7.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

### 2020-04-10 writeup-paper/old-drafts/step1_merge_data 2.Rmd
add writeup folder for apa draft

### Merge in frame times, postures, and orientations
```{r}
# calls helper functions in helper.r to get times and posture coding integrated
d <- alldets %>%
  select(-c(is_face,video)) %>% # redudant with faceMT & subid, dropping
  left_join(demo.data) %>% # join demographics info 
  group_by(subid) %>% # get subid
  do(add.times(.)) %>%  # calls helper functions to add times
  group_by(subid) %>%
  do(add.posture(.)) ## adds both postures and orientatinos
```

>>>>

### Merge in frame times, postures, and orientations
```{r}
# calls helper functions in helper.r to get times and posture coding integrated
d <- alldets %>%
  select(-c(is_face,video)) %>% # redudant with faceMT & subid, dropping
  left_join(demo.data) %>% # join demographics info 
  group_by(subid) %>% # get subid
  do(add.times(.)) %>%  # calls helper functions to add times
  group_by(subid) %>%
  do(add.posture(.)) ## adds both postures and orientatinos
```
### Recheck posture syncing times
```{r}
sync_times_check <- read_csv("../data/manual_sync_times/video_sync_times.csv") 
s <- sync_times_check %>%
  rename(subid = sid) %>%
  group_by(subid) %>%
  mutate(sync_time_string =toString(sync_time_stamp)) %>%
  mutate(frames = str_split_fixed(sync_time_string,":",3)[,3]) %>%
  mutate(sec = str_split_fixed(sync_time_string,":",3)[,2]) %>%
  mutate(min = str_split_fixed(sync_time_string,":",3)[,1]) %>%
  mutate(sync_time_ms = as.numeric(min)*6000 + as.numeric(sec)*1000 + as.numeric(frames)*33.33)

posture.sync <- function(x) {
  fname <- paste("../data/posture/",
                 x$subid[1],
                 ".csv",
                 sep="")
   if (file.exists(fname)) {

    postures <- read_csv(fname)
    postures <- postures[order(postures$start),] ## add line to sort by time in case weird coding order -- the case for some kids
    x$posture_start_time = postures$start[1]
    print(paste0("**** ", x$subid[1], " has time",x$posture_start_time))
   } else {
     print(paste0("**** ", x$subid[1], " is missing posture ****"))
   }
  return(x)
}

s <- s %>%
  group_by(subid) %>%
  do(posture.sync(.))

## problematic kids
issues = (s$posture_start_time<s$sync_time_ms)
diffs =  s$posture_start_time-s$sync_time_ms
off_by = diffs[issues]
seconds_off = round(off_by / 1000*-1,2)
```
